Estimation of the road traffic sound levels based on Non-Negative Matrix Factorization

### 2017-12-19 article/articleApplied.tex
Following of the results part

\begin{figure*}
    \centering
    \subfigure[]{\label{fig:blocDiagramNMF}
    \includegraphics[width=.45\linewidth]{../image/bloc_diagram_NMF_EN.pdf}}
    \subfigure[]{\label{fig:blocDiagramFilter}	\includegraphics[width=.45\linewidth]{../image/bloc_diagram_filtrage_EN.pdf}}
    \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}.}
    \label{fig:blocDiagram}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.9\linewidth]{../image/schema_introduction_nmf.pdf}
\caption{Example of a simple NMF  for a musical content \cite{bertin_les_2009}}
\label{fig:example_NMF}
\end{figure}

The choice of the dimensions should be such that $F\times K + K \times N < F \times N$. To estimate the quality of the approximation, an objective function is used

@@

One of the main issue with the supervised approach is

To better take account for the diverse nature of urban scenes, semi-supervised NMF can be useful as it has been proposed \cite{lee_semi-supervised_2010} to offer more flexibility. This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of spectrum representative of road traffic and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated. Here, $J = 2$. The aim is to include in $\mathbf{W_r}$ the element that are not related with the traffic. The problem (\ref{eq:nmf}) become

@@

%This approach has been used TO COMPLETE.

@@

A last approach is tested based on unsupervised NMF. Usually, $\mathbf{W}$ is learnt with the help of a learning corpus by initiated it randomly. Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain . Then NMF is performed where $\mathbf{W}$ (eq. \ref{eq:updateW_unsup}) and $\mathbf{H}$ (eq.  \ref{eq:updateH_Sup}) are updated iteratively.

@@

$\cos \theta = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $\cos \theta = 0$ means that the elements are diametrically opposed. This measure allows a normalized and an invariant scale estimation of the similarity. It is then sort in descending order. The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method (Figure \ref{fig:W_THC_NMF}). It is defined as :

@@

\begin{figure}[hbtp]
\centering
\includegraphics[width=0.8\linewidth]{../image/distanceCosLinDisplay.pdf}
\caption{Example of the $\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
\label{fig:W_THC_NMF}
\end{figure}

Other thresholding methods as the \textit{soft} and the \textit{firm} and multiples way to display the distance through a sigmoïd or a Radial Basis Function have been investigated. A fast parametric study has revealed that the \textit{hard} tresholding method  with a linear representation of the similarity according to $K$ (see fig. \ref{fig:W_THC_NMF}) was the best way to get better performances.

@@

The experiment consists in estimating the traffic road sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) and for 5 $TIR$ ([-12 -6 0 6 12] dB). First, the spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap and a number of point $nfft = 2^{12}$. Therefore, the dimensions of $V$ are $F$ = 2049 and $N$ = 664.

@@

\subsubsection{Dictionary building}

The dictionary is built from a second sound database dedicated specifically to this task. It is composed of 53 audio files of passing cars. These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{simScene} database but with two different cars (). First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$). Then time/frequency windows with $w_t \times F$ dimension are applied without overlapping on the spectrogram in order to consider several spectrum for each audio file. $w_t$ is fixed at $wt = [0.5 1]$ second. In each window, the root mean square value is calculated on each frequency bin to reduce the different spectrum in one spectra. Since the number of elements given by processing all the sound database is high, in order to reduce the computational and delete redundant information, a $K$-means clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$. A special case is added where the root mean square of \textit{all} the spectrogram is applied. Each audio file generates one element $k$ of $\mathbf{W}$. An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} on 3 seconds extract of a spectrogram of a car passage (fig. \ref{fig:specW}). In the case where $w_t$ = 1 second , 3 elements are therefore extracted of the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element (fig. \ref{fig:ElementW}).

\begin{figure}
    \centering
\begin{minipage}[t]{\linewidth}
    \subfigure[]{\label{fig:specW}
    \includegraphics[width=.45\linewidth]{../image/extractionDictionary3Example.pdf}}
    \subfigure[]{\label{fig:ElementW}
    \includegraphics[width=.45\linewidth]{../image/extractionDictionary4Example.pdf}}

    \caption{Example of the creation of the dictionary from the spectrogram of a three seconds sample of car passage with two temporal windows ($w_t$ = 1 and $w_t = all$) expressed in a logarithmic scale.}
    \label{fig:dictionaryExtraction}
    \end{minipage}
\end{figure}

@@

Supervised and semi-supervised NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction. THC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate to much from the initial dictionary. The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands). Theses two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.  Furthermore, in the case of the linear frequency scale and for supervised and semi-supervised NMF, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins. Nevertheless, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is done with the original dictionary $\mathbf{W}$, as

@@

With the third octave frequency scale, only the case $f_c = 20$ kHz is applied. Finally, for the THC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step. Table \ref{tab:estimation_factors} summarizes the parameters and the related values.

\begin{table*}[]
\centering
\begin{tabular}{lcccccc}
Parameter &  \multicolumn{6}{c}{value}\\ \hline
$\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
\begin{tabular}[l]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} & \multicolumn{3}{c}{linear} & \multicolumn{3}{c}{third octave}\\ \hline
$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
\textbf{method} & \multicolumn{2}{c}{filter} & supervised NMF & semi-supervised NMF & THC NMF \\ \hline
$\mathbf{t}$ & \multicolumn{6}{c}{0.20:0.01:0.70} \\ \hline
\end{tabular}
\caption{Summary of the different parameters taken into account in the frequency low-pass filter and NMF process and their values for the estimation of the traffic sound level}
\label{tab:estimation_factors}
\end{table*}

@@

\begin{table*}[]
\centering
\begin{tabular}{cccccccc}
$\mathbf{K}$ & $\mathbf{w_t}$ (s) & $\mathbf{f_c}$ (kHz) &\begin{tabular}[c]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} &  $\mathbf{\beta}$ & \textbf{method} & $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \hline
 &  & 20 &  &  & filter &  & 4.69 ($\pm$ 4.52) \\ 
 &  & 0.5 &  &  & filter &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
 & & 2 & spectra & 1 & supervised NMF &  & \\ 
50 & 0.5 & 0.5 & spectra & 2 & supervised NMF &  & \textbf{2.56 ($\pm$ 2.48)} \\ 
50 & 0.5 &  20 & third octave & 1 & supervised NMF &  & 3.44 ($\pm$ 3.70) \\ 
50 & 0.5 & 20 & third octave & 2 & supervised NMF &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
100 & 0.5 &  20 & spectra & 1 & semi-supervised NMF &  & \textbf{2.36 ($\pm$ 1.23)} \\ 
100 & 0.5 & 20 & spectra & 2 & semi-supervised NMF &  & \textbf{2.34 ($\pm$ 1.35)} \\ 
100 & 0.5  & 20 & third octave & 1 & semi-supervised NMF &  & \textbf{2.38 ($\pm$ 1.26)} \\ 
100 & 0.5 & 20 & third octave & 2 & semi-supervised NMF &  & \textbf{2.43 ($\pm$ 1.43)} \\ \hline \hline
 &  &  & spectra & 1 & THC NMF &  &  \\ 
100 & all & 20 & spectra & 2 & THC NMF & 0.37 & \textbf{2.22 ($\pm$ 2.37)} \\ 
100 & all & 20 & third octave & 1 & THC NMF & 0.57 & \textbf{2.19 ($\pm$ 2.01)} \\ 
100 & all & 20 & third octave & 2 & THC NMF & 0.54 & \textbf{2.16 ($\pm$ 2.24)}
\end{tabular}
\caption{Best results according to $\beta$, method and  spectral representation}
\label{tab:results}
\end{table*}

The first line resumes the error produced by the filter. $f_c = 20 $kHz is equivalent to consider all the sound source of the sound mixtures without distinguish the traffic from the others sound sources. $f_c = 500$ Hz is the cut-off frequency with the lower error. It is then the baseline to refer to compare the performance of NMF.

Then the different version of NMF is summarized with, in bold, the NMF errors lower than the baseline. With the exception of supervised NMF with third octave spectral representation, all NMF have a lower mean error than the baseline. However, the best combination is got with a THC NMF in a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54. As the standard deviation is lower, the case where $\beta = 1$ is selected too.

For each method, the best scenario is detailed according the sub-classes and the $TIR$ (Table \ref{tab:filtre_error}, \ref{tab:Sup_error}, \ref{tab:Semi_error} and \ref{tab:THC_error}).

\begin{table*}[]
\centering
\begin{tabular}{ccccccccc}
\textbf{method} & $\mathbf{\beta}$ & \textbf{\begin{tabular}[c]{@{}c@{}}spectral \\ representation\end{tabular}} & $\mathbf{f_c}$ (kHz) & \textbf{-12} & \textbf{-6} & \textbf{0} & \textbf{6} & \textbf{12} \\ \hline
filter &  &  & 20 &  &  &  &  &  \\ \hline
filter &  &  & 0.5 &  &  &  &  &  \\ \hline
Supervised NMF &  &  &  &  &  &  &  &  \\ \hline
Semi-supervised NMF &  &  &  &  &  &  &  &  \\ \hline
THC NMF &  &  &  &  &  &  &  & 
\end{tabular}
\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method}
\label{ta:results_TIR}
\end{table*}

\begin{figure}[hbtp]
\centering
\includegraphics[width=\linewidth]{../image/filter_bar.pdf}
\caption{MAE error for the best cut-off frequency for the filter method according each sub-classes and $TIR$}
\label{tab:filtre_error}
\end{figure}

\begin{figure}[hbtp]
\centering

\caption{MAE error for the best combination parameters for supervised NMF ($\beta = 2$, spectra) according each sub-classes and $TIR$}
\label{tab:Sup_error}
\end{figure}

\begin{figure}[hbtp]
\centering
\includegraphics[width=\linewidth]{../image/semi-sup_bar.pdf}
\caption{MAE error for the best combination parameters for semi-supervised NMF ($\beta = 2$, spectra) according each sub-classes and $TIR$}
\label{tab:Semi_error}
\end{figure}

\begin{figure}[hbtp]
\centering
\includegraphics[width=\linewidth]{../image/THC_bar.pdf}
\caption{MAE error for the best combination of parameters for THC NMF ($\beta = 2$, third octave) according each sub-classes and $TIR$}
\label{tab:THC_error}
\end{figure}

The first approach consider is the frequency low-pass filter. With $f_c = 20 kHz$, the error is naturally higher for low $TIR$ (-12,-6) as the traffic is less present. In the opposite, for high $TIR$ (6 and 12), as the traffic become preponderant, the traffic  sound level estimation is better as this method conserve all the traffic energy. A second cut-off frequency, $f_c = 500 Hz$, allow to have a balance as it put aside the high frequencies and to keep low frequencies. It results that for low $TIR$ with high frequencies class sound (\textit{alert}, \textit{animals}), the errors can be low but are high as soon as sound classes are composed of low frequencies (\textit{climate}, \textit{humans}, \textit{transport} and \textit{mechanics}).

Semi-supervised NMF is efficient particularly for low $TIR$ as the mobile part, $W_r$, allows to take into account the other predominant sound sources while supervised NMF is constrained to use traffic elements to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$. Nevertheless, these degrees of freedom are restrictive for high $TIR$. Indeed, semi-supervised NMF is free to include traffic components in $W_r$. The traffic sound level estimation is then less right. In the opposite, for $TIR = [6 , 12]$, supervised NMF has to use the traffic element present in $W$. In this case, the performances are then better.

Finally, THC NMF with a threshold fixed at $t = 0.54$ is the one where for low and high $TIR$, the performance are lower than the baseline's. The dictionary update allows to the method to adapt to the different $TIR$ and to only keep the closest elements of the \textit{traffic} component \ref{fig:dist_-12_12}.

@@

%\subsection{Frequency low-pass filter results}
%In a first time, the frequency low-pass filter estimator is performed on all the scene. The table \ref{tab:results_filter} summarized the mean error on the totality of the 750 scenes to find the most efficient cut-off frequency. \\
%
%\begin{table}[h]
%\centering
%\begin{tabular}{llll}
%$f_c$ (Hz) & $MAE$ (dB) & $nRMSE$  \\ \hline
% 500 & \textbf{\textcolor{red}{5.66 $\pm$6.59}} & \textbf{\textcolor{red}{1.48 $\pm$1.10}} \\
% 1000 & \textbf{6.31 $\pm$7.64} & \textbf{1.52 $\pm$1.34} \\
% 2000 & \textbf{6.65 $\pm$8.24} & \textbf{1.57 $\pm$1.54} \\
% 5000 & \textbf{7.42 $\pm$8.90} & 1.82 $\pm$1.65 \\
%10000 & \textbf{7.55 $\pm$9.00} & 1.89 $\pm$1.70 \\
%20000 & \textbf{7.59 $\pm$9.02} & 1.89 $\pm$1.72 \\
%\end{tabular}
%\caption{RMSE error for the low pass filter averaged on all the TIR and sub-classes}
%\label{tab:results_filter}
%\end{table}
%
%According the table \ref{tab:results_filter}, the cut-off frequency at 500 Hz is the most efficient on all the TIR and all the sub-classes. The corresponding errors according to the sub-classes and the TIR are summarized in the figure \ref{fig:filterAmbiance}.\\
%
%\begin{figure}[t]
%\centering

>>>>

%\begin{figure*}
%    \centering
%    \subfigure[]{\label{fig:blocDiagramNMF}
%    \includegraphics[width=.45\linewidth]{../image/bloc_diagram_NMF_EN.pdf}}
%    \subfigure[]{\label{fig:blocDiagramFilter}	\includegraphics[width=.45\linewidth]{../image/bloc_diagram_filtrage_EN.pdf}}
%    \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}.}
%    \label{fig:blocDiagram}
%\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.9\linewidth]{../image/schema_introduction_nmf.pdf}
\caption{Example of a simple NMF  for urban sound mixture, $\mathbf{W}$ and $\mathbf{V}$ are composed of 3 elements (car passages, car horn and bird's whitles)}
\label{fig:example_NMF}
\end{figure}

The choice of the dimensions is often made as that $F\times K + K \times N < F \times N$. NMF is then considered as a low rank approximation method. However, this constraint is not essential. To estimate the quality of the approximation, an objective function is used

@@

One of the main issue with the supervised approach is the generalization issue: how to be adapted to different sound mixtures with a fixed dictionary ? To better take account for the diverse nature of urban scenes, semi-supervised NMF can be useful as it has been proposed \cite{lee_semi-supervised_2010} to offer more flexibility. This method consists in composing the \textit{dictionary} with a fixed part $\mathbf{W_s} \in \mathbb{R}^+_{F\times K}$, composed in our case of spectrum representative of road traffic and with a mobile part, $\mathbf{W_r} \in \mathbb{R}^+_{F\times J}$ with $J <<K$, that is updated. Here, $J = 2$. The aim is to include in $\mathbf{W_r}$ the element that are not related with the traffic. The problem (\ref{eq:nmf}) become

@@

@@

A last approach is tested based on unsupervised NMF. Usually, $\mathbf{W}$ is learnt with the help of a learning corpus by initiated it randomly. Here, as the concerned sound source is known and audio samples of car passages are available, a initial dictionary, $\mathbf{W_0}$, is learnt by converting the audio files in the spectra domain (see part \ref{part:dictionary_learning}). Then NMF is performed where $\mathbf{W}$ (eq. \ref{eq:updateW_unsup}) and $\mathbf{H}$ (eq.  \ref{eq:updateH_Sup}) are updated alternatively.

@@

$\cos \theta = 1$ means that the elements are identical (the $k$-th element of $\mathbf{W}$ is then considered as traffic element) whereas $\cos \theta = 0$ means that the elements are significantly different. This measure allows a bound between 1 and 0 and is an invariant scale estimation of the similarity. Then, the similarity is sorted in descending order. The elements in $\mathbf{W}$ that can belong to $\mathbf{W}_{traffic}$ are then selected by a \textit{hard thresholding} method (Figure \ref{fig:W_ThC_NMF}). It is defined as :

@@

\begin{figure}[hbtp]
\centering
\includegraphics[width=0.8\linewidth]{../image/distanceCosLinDisplay.pdf}
\caption{Example of the $\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
\label{fig:W_ThC_NMF}
\end{figure}

Other thresholding methods as the \textit{soft} \cite{donoho1995noising}and the \textit{firm} \cite{fornasier2008iterative} and multiples way to display the distance through a sigmoïd or a Radial Basis Function have been investigated. A fast parametric study has revealed that the \textit{hard} tresholding method  with a linear representation of the similarity according to $K$ (see fig. \ref{fig:W_ThC_NMF}) was the best way to get better performances.

@@

The experiment consists in estimating the traffic road sound level of the 6 environmental sound sub-corpus (\textit{alert} (al), \textit{animals} (an), \textit{humans} (hu), \textit{climate} (cl), \textit{mechanics} (me), \textit{transportation} (tr)) and for 5 $TIR$ ([-12 -6 0 6 12] dB). First, the spectrogram $\mathbf{V}$ of each sound scene is built with a window size $w = 2^{12}$ with a 50 $\%$ overlap and a number of point $nfft = 2^{12}$. Therefore, the dimensions of $\mathbf{V}$ are $F$ = 2049 and $N$ = 664.

@@

\subsubsection{Dictionary building}\label{part:dictionary_learning}

The dictionary is built from a second sound database dedicated specifically to this task. It is composed of 53 audio files of passing cars. These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{SimScene} database but with two different cars (). First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$). Then time/frequency windows with $w_t \times F$ dimension are applied without overlapping on the spectrogram in order to consider several spectrum for each audio file. $w_t$ is fixed at $wt = [0.5 1]$ second. In each window, the root mean square value is calculated on each frequency bin to reduce the different spectrum in one spectra. Since the number of elements given by processing all the sound database is high, in order to reduce the computational and delete redundant information, a $K$-means clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$. A special case is added where the root mean square of \textit{all} the spectrogram is applied. Each audio file generates one element $k$ of $\mathbf{W}$. An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} on 3 seconds extract of a spectrogram of a car passage (fig. \ref{fig:specW}). In the case where $w_t$ = 1 second , 3 elements are therefore extracted of the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element (fig. \ref{fig:ElementW}).

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.47\linewidth}
        \centering
       \includegraphics[width=\linewidth]{../image/dictionary1.pdf}
        \caption{Sample of 3 seconds of a car passage}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.47\linewidth}
        \centering
       \includegraphics[width=\linewidth]{../image/dictionary2.pdf}
        \caption{Elements of the dictionary got for $w_t = 1$ second (3 spectrum) and  for $w_t = all$ (1 spectra) }
        \label{fig:TIR_class_filter}
    \end{subfigure}
\end{figure}

@@

Supervised and semi-supervised NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction. ThC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate to much from the initial dictionary. The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands). Theses two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.  Furthermore, in the case of the linear frequency scale and for supervised and semi-supervised NMF, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins. Nevertheless, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is done with the original dictionary $\mathbf{W}$, as

@@

With the third octave frequency scale, only the case $f_c = 20$ kHz is applied. Finally, for the ThC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step. Table \ref{tab:estimation_factors} summarizes the parameters and the related values.

\begin{table*}[]
\centering
\begin{tabular}{lcccccc}
Parameter &  \multicolumn{6}{c}{value}\\ \hline
$\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
\begin{tabular}[l]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} & \multicolumn{3}{c}{linear} & \multicolumn{3}{c}{third octave}\\ \hline
$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
\textbf{method} & \multicolumn{2}{c}{filter} & supervised NMF & semi-supervised NMF & ThC NMF \\ \hline
$\mathbf{t}$ & \multicolumn{6}{c}{0.20:0.01:0.70} \\ \hline
\end{tabular}
\caption{Summary of the different parameters taken into account in the frequency low-pass filter and NMF process and their values for the estimation of the traffic sound level}
\label{tab:estimation_factors}
\end{table*}

@@

\begin{table*}[]
\centering
\begin{tabular}{@{}cccccccc@{}}
\toprule
$\mathbf{K}$ & $\mathbf{w_t}$ (s) & $\mathbf{f_c}$ (kHz) &\begin{tabular}[c]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} &  $\mathbf{\beta}$ & \textbf{method} & $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
 &  & 20 &  &  & filter &  & 4.69 ($\pm$ 4.52) \\ 
 &  & 0.5 &  &  & filter &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
25 & all & 0.5 & spectra & 1 & supervised NMF &  & \textbf{2.76 ($\pm$ 2.86)} \\ 
50 & 0.5 & 0.5 & spectra & 2 & supervised NMF &  & \textbf{2.56 ($\pm$ 2.48)} \\ 
50 & 0.5 &  20 & third octave & 1 & supervised NMF &  & 3.44 ($\pm$ 3.70) \\ 
50 & 0.5 & 20 & third octave & 2 & supervised NMF &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
100 & 0.5 &  20 & spectra & 1 & semi-supervised NMF &  & \textbf{2.36 ($\pm$ 1.23)} \\ 
100 & 0.5 & 20 & spectra & 2 & semi-supervised NMF &  & \textbf{2.34 ($\pm$ 1.35)} \\ 
100 & 0.5  & 20 & third octave & 1 & semi-supervised NMF &  & \textbf{2.38 ($\pm$ 1.26)} \\ 
100 & 0.5 & 20 & third octave & 2 & semi-supervised NMF &  & \textbf{2.43 ($\pm$ 1.43)} \\ \hline \hline
 &  &  & spectra & 1 & ThC NMF &  &  \\ 
100 & all & 20 & spectra & 2 & ThC NMF & 0.37 & \textbf{2.22 ($\pm$ 2.37)} \\ 
100 & all & 20 & third octave & 1 & ThC NMF & 0.57 & \textbf{2.19 ($\pm$ 2.01)} \\ 
100 & all & 20 & third octave & 2 & ThC NMF & 0.54 & \textbf{2.16 ($\pm$ 2.24)}\\ \bottomrule
\end{tabular}
\caption{Best results according to $\beta$, method and  spectral representation}
\label{tab:results}
\end{table*}

The two first lines resume the error produced by the filter. $f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinguish the traffic from the others sound sources. Consequently , in low $TIR$ (-12 and -6), where traffic component is discreet, the error is more important than in high $TIR$ (6 and 12) where the traffic component is predominant. $f_c = 500$ Hz is the cut-off frequency with the lower mean error. It is then the baseline to refer to compare the performance of NMF. In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, the impact of the filter is quite interesting as it suppress theses frequency components whereas for the other sub-classes where low frequencies are presents, the error as higher as the filter cannot dissociate the traffic element from the other sound sources. The sounds levels are then overestimate (figure \ref{fig:alert_-12}). In opposite, in high $TIR$, the low-pass filter removes too much energy from the traffic which has the consequence to underestimate the sound levels (figure \ref{fig:alert_12}).\\

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/LeqTrafficComparison_filter_alert_-12.pdf}
        \caption{}
        \label{fig:alert_-12}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/LeqTrafficComparison_filter_alert_12.pdf}
        \caption{}
        \label{fig:alert_12}
    \end{subfigure}

    \caption{Global sound levels of the traffic estimated by the frequency low-pass filter with $f_c$ = 500 Hz for the sub-classes \textit{alert}: at $TIR = -12$ (\ref{fig:alert_-12}) and at $TIR$ (\ref{fig:alert_12}).}
    \label{fig:dictionaryExtraction}
\end{figure*}

The results of different versions of NMF are summarized next to the filter results with, in bold, the NMF errors lower than the baseline. With the exception of supervised NMF with third octave spectral representation, all NMF have a lower mean error than the baseline. However, the best combination is got with a ThC NMF with a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54. For each method (supervised, semi-supervised and ThC NMF), the best scenario is detailed according the sub-classes and the $TIR$ (Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:W_ThC_NMF}).

\begin{table*}[t]
\centering
\begin{tabular}{@{}cccccc@{}}
\toprule
method & filter & filter & supervised NMF & semi-supervised NMF & ThC NMF \\ \midrule
$\beta$ &  &  & 2 & 2 & 2 \\
\begin{tabular}[c]{@{}c@{}}spectral \\ representation\end{tabular} &  &  & spectra & spectra & third octave \\
$f_c$ (kHz) & 20 & 0.5 & 0.5 & 20 & 20 \\ \hline
-12 & 12.25 ($\pm$ 0.05) & 7.36 ($\pm$ 3.00) & \textbf{6.23 ($\pm$ 3.19)} &  & \textbf{5.11 ($\pm$ 3.10)} \\
-6 & 6.96 ($\pm$ 0.05) & 3.44 ($\pm$ 1.65) & \textbf{3.00 ($\pm$ 1.39)} &  & \textbf{2.87 ($\pm$ 1.55)} \\
0 & 3.00 ($\pm$ 0.03) & 1.17 ($\pm$ 0.24) & \textbf{1.14 ($\pm$ 0.17)} &  & 1.38 ($\pm$ 0.38) \\
6 & 0.97 ($\pm$ 0.01) & 1.03 ($\pm$ 0.26) & \textbf{1.01 ($\pm$ 0.28)} &  & \textbf{0.70 ($\pm$ 0.32)} \\
12 & 0.26 ($\pm$ 0.00) & 1.45 ($\pm$ 0.13) & \textbf{1.41 ($\pm$ 0.10)} &  & \textbf{0.76 ($\pm$ 0.19)} \\ \bottomrule
\end{tabular}
\caption{$MAE$ error averaged on all sub-classes on each $TIR$ for the best scenario according to each method}
\label{ta:results_TIR}
\end{table*}

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/filter_bar.pdf}
        \caption{filter method with $f_c$ = 500 Hz}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/sup_bar.pdf}
        \caption{Supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 0.5 kHz}
                \label{fig:TIR_class_sup}
    \end{subfigure}

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/semi-sup_bar.pdf}
        \caption{Semi-supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 20 kHz}
                \label{fig:TIR_class_semi}
    \end{subfigure}%
    \hfill 
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/THC_bar}
        \caption{ThC NMF, $\beta$ = 2, third octave spectral representation, $f_c$ = 20 kHz}
        \label{fig:TIR_class_thc}
    \end{subfigure}
    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter and each method (supervised, semi-supervised and ThC}
\end{figure*}

In the case of supervised NMF, on all the $TIR$, if the mean score is lower than the baseline method, it offers similar performances as for $TIR = [0, 6, 12]$, the mean score are nearly close. The performance of supervised NMF is mainly visible for $TIR = -12$ where the error with the baseline decrease by more than 1 dB.

The gain of the semi-supervised approach is most of all significant for these low $TIR$ where the add of a mobile part, $\mathbf{W_r}$, into $\mathbf{W}$ allows to take into account the other predominant sound sources (example in Figure \ref{fig:Y_alert-12}) while supervised NMF is constrained to use traffic elements to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$. This improvement, according to supervised NMF results, can be seen in Figure \ref{fig:TIR_class_semi} where the error on each sub class for $TIR = -12$ and $TIR$ = -6 are much lower.  Nevertheless, these degrees of freedom are restrictive for high $TIR$. Indeed, without constraint semi-supervised NMF is free to include traffic components in $W_r$ decreasing the rebuilt of the traffic component (Figure \ref{fig:Y_alert_12}). The error are then increasing for theses $TIR$. In the opposite, supervised NMF has to use the \textit{traffic} element present in $W$ which improve the performances.

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_-12_01.pdf}
        \caption{$\mathbf{W_r}$ for an \textit{alert} scene at $TIR$ = -12}
        \label{fig:Y_alert-12}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_12_01.pdf}
        \caption{$\mathbf{W_r}$ for an \textit{alert} scene at $TIR$ = 12}
		\label{fig:Y_alert_12}
    \end{subfigure}
\end{figure}

Finally, ThC NMF with a threshold fixed at $t = 0.54$ offers the most lower results. The dictionary update allows to the method to adapt to the different $TIR$ and to only keep the closest elements of the \textit{traffic} component \ref{fig:dist_-12_12}.

@@

The gain is significant on all $TIR$ with the exception of $TIR = 0$. On figure \ref{fig:ThC_error}, we can see the improvement of .
While supervised NMF have high errors on low $TIR$ for the sub-classes \textit{human} and \textit{transport}, ThC NMF succeed to have error inferior to 4 dB. The error for \textit{climate} and \textit{mechanics} are, on the other hand, similar to those of supervised NMF.
